Concise Linkable Ring Signatures and Applications

### 2020-01-06 paper/iacr.tex
Paper update for IACR archive
Paper updates

\documentclass[draft]{article} % TODO: remove draft
\usepackage[top=1.0in,bottom=1.0in,left=1.0in,right=1.0in]{geometry}
\usepackage{amsmath,amssymb,amsthm,amsfonts}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{graphicx}

% TODO remove
\usepackage{lineno}
\linenumbers

\title{Triptych: logarithmic-sized linkable ring signatures with applications}
\author{Sarang Noether \\ Monero Research Lab \\ \texttt{sarang.noether@protonmail.com}}
\date{\today}

\newcommand{\G}{\mathbb{G}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\com}{\operatorname{Com}}
\newcommand{\hs}{\mathcal{H}}
\newcommand{\hp}{\mathbb{H}}

\newcommand{\A}{\mathcal{A}}

\newcommand{\sumi}{\sum_{i=0}^{n-1}}
\newcommand{\sumj}{\sum_{j=0}^{m-1}}
\newcommand{\sumk}{\sum_{k=0}^{N-1}}

\newtheorem{theorem}{Theorem}
\theoremstyle{definition}
\newtheorem{definition}{Definition}

\begin{document}

\maketitle

\begin{abstract}
Ring signatures are a common construction used to provide signer ambiguity among a non-interactive set of public keys specified at the time of signing.
Unlike early approaches where signature size is linear in the size of the signer anonymity set, current optimal solutions either require centralized trusted setups or produce signatures logarithmic in size.
However, few also provide linkability, a property used to determine whether the signer of a message has signed any previous message, possibly with restrictions on the anonymity set choice.
Here we introduce Triptych, a family of linkable ring signatures without trusted setup that is based on generalizations of zero-knowledge proofs of knowledge of commitment openings to zero.
We demonstrate applications of Triptych in signer-ambiguous transaction protocols by extending the construction to openings of parallel commitments in independent anonymity sets.
Signatures are logarithmic in the anonymity set size and, while verification complexity is linear, collections of proofs can be efficiently verified in batches.
We show that for anonymity set sizes practical for use in distributed protocols, Triptych offers competitive performance with a straightforward construction.
\end{abstract}

\section{Introduction}
First introduced in \cite{rivest} with respect to RSA groups, ring signatures permit the signing of messages using public key sets not fixed in advance, without the need of a trusted group manager.
Earlier constructions lacked such flexibility, requiring either centralized key setup or the establishment of fixed signing sets.
Later work \cite{bender} established more robust security models for unforgeability and anonymity, capturing realistic threat models where an adversary is permitted to corrupt keys, convince honest signers to include malicious anonymity set members, or obtain signatures in advance.

Since a ring signature has an anonymity set of public keys, one of which is the true signer, the detection of signing by the same key requires an additional property, linkability.
A linkable ring signature \cite{lsag} enables verifiers to determine whether the (unknown) signer of a message has signed other messages.
Such a construction was proposed for election applications, where it is necessary to ensure that votes are anonymous, but voters are permitted to vote only once on a particular issue.
The construction in \cite{lsag}, with a hash-trapdoor structure similar to that of \cite{schnorr}, is of particular interest due to potential flexibility in linking; while its linking is limited to signer-selected but unchanging signer anonymity sets, it permits linking on a per-issue basis.
Additional recent work in \cite{backes} introduces the property of linkable anonymity, using sets of signatures and establishing restrictions on key corruption.
Other related interesting properties like traceability \cite{fs,f} imply stronger capabilities, where an attempt to sign two messages with the same key permits verifiers to identify the signer.

Linkable ring signatures have seen particular applications in signer-ambiguous transaction protocols.
In such an application, transactions are authorized by a ring signature whose signing anonymity set consists of previously-generated transaction outputs.
A signature demonstrates that the signer controls the private key of one such output without revealing which is the signer, and linkability is used to assure verifiers that output has not been used in another signature (signifying a double-spend attempt).

A practical consideration for the use of linkable ring signatures in transaction protocols is how signature size and verification time scale with the size of the anonymity set.
In common applied constructions like \cite{mlsag,clsag}, signature size and verification time scale linearly with the size of the signing ambiguity set; since such signatures are typically included in a public distributed data structure like a blockchain, there is a balance between the size of the anonymity set and the requirements for storage and verification.
Recent protocol work in this area mitigates the size restriction.
For example, in \cite{rct3}, the authors introduce a confidential transaction protocol based on a proving system whose size scales logarithmically with the anonymity set size, and which includes a method for demonstrating amount balance; amount commitment range proving is offloaded to other constructions like \cite{bulletproofs}.
In \cite{omniring}, the authors use a more general proving technique to accomplish a similar goal; however, the protocol offers further size benefits by integrating commitment range proofs into the proof structure directly, taking advantage of the logarithmic proof size.

Other signer-ambiguous transaction protocols not based on linkable ring signatures offer more competitive performance.
For example, protocols like \cite{zcash} produce maximal theoretical signer ambiguity through the use of zero-knowledge Merkle proofs (among others) that offer extremely small proofs with low verification time, but at the cost of a trusted structured setup process that arises from the underlying proving system \cite{groth16}.
Like this work, \cite{lelantus} is a transaction protocol also based on \cite{groth}; however, it is intended to operate on commitments similar to those used in \cite{zerocoin} but with the inclusion of amounts, and has limitations on addressing and sender tracing.

\subsection{Our contribution}
We produce two instantiations of a linkable ring signature, which we call Triptych.
The constructions are a linkable generalization of Groth's one-of-many commitment-to-zero proving system \cite{groth}, with optimizations from Bootle \cite{bootle} applied for improvements to proof size and verification complexity.
In one version of Triptych, the prover shows that it knows the opening of a commitment to zero within a commitment list, and also that it has constructed a linking tag using the same opening, enabling linkability.
We then modify Groth's ring signature definitions to include linkability and a related property, non-frameability.

In the second version of Triptych, we further generalize to include a parallel list of commitments.
Here, the prover proceeds as before to show its knowledge of an opening to a commitment in one list, as well as the construction of the linking tag.
However, the proof also shows that the prover knows an opening of a commitment at the same position in the second commitment list.
This construction has immediate application; in some signer-ambiguous confidential transaction protocols, transaction inputs are commitments to zero, for which the signer shows it knows the opening.
Each commitment comes equipped with another commitment to the amount of the input; by offsetting these commitments homomorphically and carefully choosing commitment randomness, the prover can show that a particular transaction balances.

We show that Triptych produces signatures with competitive performance to other modern linkable ring signatures for limited anonymity set sizes.
We note that similar constructions also require linear verification time, meaning that the size of anonymity set used in practice is likely to be limited for performance reasons.

\section{Preliminaries}
\subsection{Public parameters}
Let $\G$ be a cyclic group in which the discrete logarithm problem is hard, and let $\F$ be the scalar field of $\G$.
Let $\hs: \{0,1\}^* \to \F$ be a cryptographic hash function.
Let $G$ and $H$ be generators of $\G$ with unknown discrete logarithm relationship.
Let $N = n^m$ be a size parameter, where $n > 1$ and $m > 1$.
Let $\{G_{j,i}\}_{j,i=0}^{m-1,n-1}$ be a set of generators of $\G$ with unknown discrete logarithm relationship to each other, to $G$, and to $H$.
Let $U$ be a generator of $\G$.
Note that all generators may be produced using public randomness; for example, the use of a suitable hash function with domain separation may be appropriate.
All such public parameters are assumed to comprise a global reference string known to all players; in particular, we exclude them from algorithm definitions and Fiat-Shamir transcript hashes for readability.

\subsection{Pedersen commitment}
Let $\com$ be a homomorphic commitment scheme that is perfectly hiding and computationally binding.
In this work, we assume use of the Pedersen commitment scheme: for $x,r \in \F$, define $\com(x,r) \equiv xG + rH$ to be the commitment of the value $x$ with randomness $r$.
This can be trivially extended to support matrix values; for $\{x_{j,i}\},r \in \F$, define $\com(x,r) \equiv rH + \sum_{j,i} x_{j,i}G_{j,i}$.
Note in particular that Pedersen matrix commitments are similarly homomorphic.

\subsection{Other notation}
For integers or field elements $i,j$, the Kronecker delta function $\delta(i,j)$ evaluates to $1$ if $i=j$ and $0$ otherwise, where the output is taken to be in the appropriate set.

We sometimes use index subscript notation of the form $i_j$ to indicate the $j$ digit of $i$, where such a decomposition of $i$ is taken base $n$ with padded length $m$:
$$\sum_{j=0}^m i_j n^j = i$$
This notation will be specified explicitly where confusion may occur.

\section{Protocol: linkable one-of-many commitment}
We wish to build a linkable ring signature construction, where a signer who knows the opening of a commitment may sign messages using an anonymity set containing other commitments for which the signer does not know openings.
Included with the proof of knowledge, the signer also provides a linking tag that is the image of the signing commitment's opening under a verifiable pseudorandom function, using the method of \cite{dodis} that has previously appeared in \cite{omniring,rct3}.
Part of the soundness of the proving system relies on the proper construction of this linking tag.
Upon receipt, a verifier can check whether the linking tag has previously appeared in any other valid proof; if it has not, injectivity assures the verifier that no other signature has been produced by the (unknown) signer.

More specifically, we modify the construction of Bootle \cite{bootle}, which itself is a generalization of a construction by Groth \cite{groth}.
We produce a sigma protocol for the following relation:
$$\mathcal{R}_{\text{link}} = \left\{ \{M_i\}_{i=0}^{N-1} \subset \G, J \in \G ; (l \in \mathbb{Z},r \in \F) : M_l = rG \text{ and } U = rJ \right\}$$
Figures \ref{fig:link} and \ref{fig:link_2} describe the protocol.

\begin{figure}[htbp]
\centering
\fbox{\begin{minipage}{0.95\textwidth}
$\underline{\mathcal{P}_{\text{link}}(\{M_i\},J;(l,r)):}$
\begin{itemize}
\item Select random $r_A \in \F$ and $\left\{a_{j,i}\right\}_{i=1,j=0}^{n-1,m-1} \subset \F$.
Define $A \equiv \com(a,r_A)$.
\item Define $\left\{\sigma_{j,i}\right\}_{i,j=0}^{n-1,j-1} \subset \F$ such that $\sigma_{j,i} \equiv \delta\left(l_j,i\right)$ (using our decomposition notation), and choose random $r_B \in \F$.
Define $B \equiv \com(\sigma,r_B)$.
\item Select random $r_C \in \F$, and define $C \equiv \com(a(1-2\sigma), r_C)$.
\item Select random $r_D \in \F$, and define $D \equiv \com(-a^2, r_D)$.
\item Define coefficients $\left\{p_{k,j}\right\}_{k,j=0}^{N-1,m-1}$ such that $$p_k(x) \equiv \prod_{j=0}^{m-1} \left( \sigma_{j,k}x + a_{j,k} \right) = \delta\left(l,k\right)x^m + \sumj p_{k,j}x^j$$ for all $k \in [0,N)$ (using our decomposition notation).
\item Select random $\left\{\rho_j\right\}_{j=0}^{m-1} \subset \F$.
\item Define $\{X_j\}_{j=0}^{m-1} \subset \G$ such that: $$X_j \equiv \sumk p_{k,j}M_k + \rho_jG$$
\item Define $\{Y_j\}_{j=0}^{m-1} \subset \G$ such that: $$Y_j \equiv U \sumk p_{k,j}+ \rho_jJ$$
\end{itemize}

$\underline{\mathcal{P} \to \mathcal{V}:}$ \\
$A,B,C,D,\{X_j\},\{Y_j\}$ \\

$\underline{\mathcal{V} \to \mathcal{P}:}$ \\
$\xi \in \{0,1\}^*$ \\

$\underline{\mathcal{P}(\xi):}$
\begin{itemize}
\item Define $\left\{f_{j,i}\right\}_{i=1,j}^{n-1,m-1}$ such that $f_{j,i} \equiv \sigma_{j,i}\xi + a_{j,i}$.
\item Define $z_A \equiv r_A + \xi r_B$ and $z_C \equiv \xi r_C + r_D$.
\item Define $z \equiv r \xi^m - \sumj \rho_j \xi^j$.
\end{itemize}

$\underline{\mathcal{P} \to \mathcal{V}:}$ \\
$\{f_{j,i}\}_{j=0,i=1}^{m-1,n-1},z_A,z_C,z$
\end{minipage}}
\caption{Sigma protocol for $\mathcal{R}_{\text{link}}$}
\label{fig:link}
\end{figure}

\begin{figure}[htbp]
\centering
\fbox{\begin{minipage}{0.95\textwidth}
$\underline{\mathcal{V}_{\text{link}}(\{M_i\},J):}$
\begin{itemize}
\item For $0 \leq j < m$, let $f_{j,0} \equiv \xi - \sum_{i=1}^{n-1} f_{j,i}$.
\item Accept if and only if:
\begin{eqnarray}
A + \xi B &=& \com(f,z_A) \label{eqn:ab} \\
\xi C + D &=& \com(f(\xi-f),z_C) \label{eqn:cd} \\
\sumk M_k \left( \prod_{j=0}^{m-1} f_{j,k_j} \right) - \sumj \xi^jX_j - zG &=& 0 \label{eqn:x} \\
U \sumk \left( \prod_{j=0}^{m-1} f_{j,k_j} \right) - \sumj \xi^jY_j - zJ &=& 0 \label{eqn:y}
\end{eqnarray}
\end{itemize}
\end{minipage}}
\caption{Sigma protocol for $\mathcal{R}_{\text{link}}$ (continued)}
\label{fig:link_2}
\end{figure}

Observe that this protocol can be made non-interactive using the Fiat-Shamir heuristic, where the verifier challenge is produced using a collision-resistant hash function (modeling a random oracle) and the proof transcript \cite{fiatshamir}.

We will show that the sigma protocol is complete, sound, and zero-knowledge, the precise definitions of which are common and found in \cite{groth}.
Informally, we require the protocol be:
\begin{itemize}
\item \textit{Perfectly complete}: Given knowledge of a witness to a statement in the proof relation, an honest prover can always convince an honest verifier of the validity of the witness.
\item \textit{Special sound}: Given a statement in the proof relation, if a prover can answer multiple honest verifier challenges correctly, then it is possible to extract a witness for this statement.
\item \textit{Special honest-verifier zero knowledge}: Given any statement and verifier challenge, it is possible to simulate a transcript that is accepted by an honest verifier without knowledge of a corresponding witness.
\end{itemize}

\begin{theorem}
The protocol in Figures \ref{fig:link} and \ref{fig:link_2} is perfectly complete, special honest-verifier zero knowledge, and $(m+1)$-special sound.
\end{theorem}

\begin{proof}
The proof follows similarly to that of \cite{bootle}.

% Completeness
We first show perfect completeness.
Suppose the verifier receives a proof generated by an honest prover.
Equation \ref{eqn:ab} holds using the identity $$\sumi \sigma_{j,i} = 1$$ for all $0 \leq j < m$.
Equation \ref{eqn:cd} follows similarly, using the identity $$\left(\sigma_{j,i}\right)^2 = \sigma_{j,i}$$ for all $0 \leq j < m$.
To show Equation \ref{eqn:x} holds:
\begin{eqnarray*}
&& \sumk M_k \left( \prod_{j=0}^{m-1} f_{j,k_j} \right) - \sumj \xi^jX_j - zG \\
&=& \sumk M_k p_k(\xi) - \sumj \xi^j \left( \sumk p_{k,j}M_k + \rho_jG \right) - zG \\
&=& \sumk M_k \left( p_k(\xi) - \sumj \xi^j p_{k,j} \right) - \sumj \xi^j \rho_jG - zG \\
&=& \sumk M_k \xi^m \delta(l,k) - \sumj \xi^j \rho_jG - \left( r\xi^m - \sumj \rho_j\xi^j \right)G \\
&=& \xi^mrG - \sumj \xi^j \rho_jG - \xi^mrG + \sumj \xi^j \rho_jG \\
&=& 0
\end{eqnarray*}
Equation \ref{eqn:y} follows similarly:
\begin{eqnarray*}
&& U \sumk \left( \prod_{j=0}^{m-1} f_{j,k_j} \right) - \sumj \xi^jY_j - zJ \\
&=& U \sumk p_k(\xi) - \sumj \xi^j \left( U \sumk p_{k,j} + \rho_jJ \right) - zJ \\
&=& U \sumk \left( p_k(\xi) - \sumj \xi^j p_{k,j} \right) - \sumj \xi^j \rho_jJ - zJ \\
&=& U \sumk \xi^m \delta(l,k) - \sumj \xi^j \rho_jJ - \left( r\xi^m - \sumj \rho_j\xi^j \right)J \\
&=& \xi^mU - \sumj \xi^j \rho_jG - \xi^mrJ + \sumj \xi^j \rho_jG \\
&=& 0
\end{eqnarray*}
since $J = r^{-1}U$ in a valid proof.
Hence the protocol is perfectly complete.

% SHVZK
We next show that the protocol is special honest-verifier zero knowledge.
To do so, we construct a simulator that, given a random verifier challenge $\xi$, can construct a proof transcript with identical distribution to a valid proof.

First, observe that the simulator presented in the proof of Lemma 1 in \cite{bootle} translates identically to our setting.
If the simulator chooses $B \in \G$ uniformly at random, the cited lemma assures us a valid simulation of the proof elements $A,C,D,z_A,z_C,\{f^{(u)}_{j,i \neq 0}\}$; we may compute each $f^{(u)}_{j,0}$ from this.
Further, in a valid proof, $B$ is independent and uniformly distributed as well.

The proof elements $\{X_j\}_{j=1}^{m-1}$ and $\{Y_j\}_{j=1}^{m-1}$ are independent and uniformly distributed in a valid proof since the set $\{\rho_j\}$ is random and the discrete logarithm problem in $\G$ is hard, so the simulator may choose these uniformly at random.
The verification checks require that $X_0$ and $Y_0$ be uniquely determined by the other elements in the corresponding sets in both real proofs and by the simulator.

Finally, $z$ is uniformly distributed in valid proofs given random $\xi$, so the simulator may choose it uniformly at random.
Hence the construction is special honest verifier zero-knowledge.

>>>>

\documentclass[draft]{llncs} % TODO: remove draft option
\pagestyle{headings}

\usepackage{enumerate}
\usepackage{amsmath,amssymb}
\usepackage{hyperref}
\usepackage[notcite]{showkeys} % TODO: remove showkeys
\usepackage{lineno} % TODO: remove
\linenumbers % TODO: remove

\title{Concise Linkable Ring Signatures and Applications}
\institute{Monero Research Lab, \email{\{surae,sarang\}@getmonero.org} \and Independent researcher, \email{randomrun@protonmail.com}}
\author{Brandon Goodell\inst{1} \and Sarang Noether\inst{1} \and Arthur Blue\inst{2}}

\begin{document}
\maketitle

\begin{abstract}
We describe a generalization of linkable ring signatures, $d$-linkable ring signatures ($d$-LRS). These admit $d$-dimensional vectors of keys with a specified linking coordinate, but do not have signature sizes linear in $d$. We then produce an efficient $d$-LRS construction, called \textit{concise linkable spontaneous anonymous group} signatures (CLSAG), that are smaller than existing constructions without a trusted setup used in production, and are more efficient in verification over a reasonable range of anonymity set sizes. Security is proven using the $k$-one more discrete logarithm and decisional Diffie-Hellman hardness assumptions to show correctness, unforgeability, linkable anonymity, linkability, and non-frameability. We describe direct applications to signer-ambiguous confidential transactions with single or multiple asset classes.
\end{abstract}

\section{Introduction}

First introduced in \cite{rivest} in the RSA setting and in \cite{liu} in the discrete logarithm setting, ring signatures permit message signing on behalf of a set of public keys rather than a single public key. Ring signatures see myriad applications ranging from lightweight anonymous authentication in \cite{yang2015lightweight} to transaction protocols like Monero in \cite{noether} and CryptoNote in \cite{van2013cryptonote}. A verifier is assured that the signer knows the private key of at least one of these public keys, which are called ring members. Ring signatures are anonymous or signer-ambiguous in the sense that the verifier does not learn information from the signature about which key is the signer. We stress that methods of practical analysis such as those of  \cite{moser2018empirical,quesnelle2017linkability} can exploit metadata in real-life applications of anonymous authentication protocols to reduce anonymity.

Group signature constructions preceding \cite{rivest,liu} require some degree of interactivity, a fixed set of participants, a trusted group manager or other trusted setup, or hardness assumptions not based on the discrete logarithm problem. Since \cite{rivest}, ring signatures have enjoyed many improvements, extensions, and modifications. For example, ring signatures are constructed in the bilinear pairing setting in \cite{zhang2002id}, key structures are generalized in \cite{abe20021}, security definitions are improved in \cite{bender}, signature size is improved in \cite{fujisaki2011sub,gu2018constant}, and traceability is introduced in \cite{fujisaki2007traceable}.

Linkable ring signatures were first introduced in \cite{liu}; in the context of transaction protocols, linkable ring signatures are the basis for anonymous transaction authentication. Linkable ring signatures guarantee that two signatures with the same ring on arbitrary messages can be publicly linked if signed using the same key. An implementation is presented in \cite{liu} in the discrete logarithm setting; that implementation functions for similar reasons as Schnorr signatures in \cite{schnorr1991efficient}.

The signature linking tags in \cite{liu} are unsuitable for applications where signatures must be linked key-by-key, not ring-by-ring (such as for ``double-signing'' protection in a setting where users generate new keys over time and select \textit{ad hoc} ring members). Resistance to double-signing attempts is ensured using linking tags as described in \cite{van2013cryptonote} (therein called key images). More recent work of \cite{noether} extends the approach of \cite{liu} to enable an anonymous confidential transaction model with \textit{ad hoc} ring member selection. In \cite{noether}, transaction amounts are replaced with Pedersen commitments to amounts together with range proofs. Signatures are constructed from key vectors including differences of amount commitments as one of the keys. However, the proofs in \cite{noether} are informal and not based on rigorous security models.

Alternatives to ring signatures like more general zero-knowledge proving systems typically require a trusted party to honestly perform a setup process (as in \cite{groth,ben2014succinct,groth2018updatable}) or lack practical efficiency for large circuits (as in \cite{bulletproofs}), meaning that such systems may not be appropriate for distributed ledger applications. However, more recent approaches such as \cite{bowehalo,hoffmann2019efficient} show both improvements to the trust requirement as well as improvements in efficiency.

\subsection{Our contribution}

We first introduce a formal definition of $d$-dimensional linkable ring signatures ($d$-LRS) as a generalization of linkable ring signatures (LRS). We present an implementation of a new $d$-LRS, $d$-concise linkable spontaneous anonymous group ($d$-CLSAG) signatures. These have $d$-dimensional keys, and signature size scales with the sum of ring size and the dimension $d$. Equivalent MLSAG signatures of \cite{noether} produce signatures that scale with the product of ring size and $d$ rather than the sum. Size efficiency comes from an aggregation of keys across key vector components, similar to the approach from \cite{maxwell2018simple,qian2010non}; this results in signatures that are about half the size of \cite{noether}, and can be generated and verified more quickly in practice.

We describe linkable ring signatures, $d$-LRS constructions, and examples in Section \ref{sec:lrs}. We present our $d$-CLSAG construction in Section \ref{sec:construction} and discuss its efficiency. The security model and main results are presented in Section \ref{sec:security}. We produce informal descriptions of signer-ambiguous confidential transaction applications for single and multiple asset types in Section \ref{sec:applications}. Security proof are presented in Appendix \ref{app:proofs}.

Notably, we provide a new definition for unforgeability in linkable ring signatures that takes into account both insider corruption and forgeries from partially-corrupted rings. We prove that $d$-CLSAG signatures are unforgeable under the hardness of the $k$-one-more discrete logarithm ($k$-OMDL) problem under this definition in the random oracle model, and show that  the properties of linkability and non-frameability follow from unforgeability. We show that linkable anonymity reduces to the hardness of a form of the decisional Diffie-Hellman (DDH) problem.

\subsection{Notation}\label{sec:notation}

We denote algorithms with typefont upper-case English letters like $\texttt{A}$, $\texttt{B}$, or $\texttt{O}$, or typefont names like $\texttt{Setup}$, $\texttt{KeyGen}$, and so on.  For any prime $p$, denote the field with $p$ elements as $\mathbb{F}_p := \mathbb{Z}/p\mathbb{Z}$, and denote the non-zero elements as $\mathbb{F}_p^*$.

Group parameters are denoted as a tuple $(p, \mathbb{G}, d, G)$ where $\mathbb{G}$ is an elliptic curve group with prime order $p$, $d$ is a dimension, and $G$ is a generator of $\mathbb{G}$. We denote integers, bits, indices, and scalars in $\mathbb{F}_p$ with lower-case English letters $x, y, z, b, c, i, j, k$, etc.\ and we denote group elements with upper-case English letters, $G, X, W$, and so on. We use lower-case Greek letters like $\sigma$ to describe signatures and upper-case calligraphic Latin letters like $\mathfrak{T}$ when describing key images.

For these group parameters, the secret key space is $\mathbb{F}_p^*$ and the public key space is $\mathbb{G}$. For any non-zero secret key $\textit{sk} \in \mathbb{F}_p^*$, the corresponding public key $\textit{pk}$ is computed from the generator $G$ through exponentiation in $\mathbb{G}$ as usual. However, we use notation for a module over the field $\mathbb{Z}_p$ to maintain consistency with related work like \cite{van2013cryptonote} and \cite{noether}.

We denote column vectors in boldface, such as $(x_1, \ldots, x_d)^\top = \textbf{x}$, and matrices in underlined boldface, such as $((x_{1,1}, x_{1,2}, \ldots, x_{1,n}), \ldots, (x_{d,1}, \ldots, x_{d,n})) = (\textbf{x}_1, \textbf{x}_2, \ldots, \textbf{x}_n) = \underline{\textbf{x}}$ is a $d\times n$ matrix. We denote the Hadamard product of two vectors with $\circ$, so for any $\textbf{x} = (x_1, x_2, \ldots, x_d) = (x_i)_{i=1}^{d}$, and for any $\textbf{y} = (y_i)_{i=1}^{d}$, we denote the sequence $(x_i \cdot y_i)_{i=1}^{d}$ with $\textbf{x} \circ \textbf{y}$. We denote bitwise concatenation with the symbol $\mid \mid$.

We distinguish oracles with calligraphic font; $\mathcal{CO}$ denotes a corruption oracle, $\mathcal{SO}$ denotes a signing oracle, and so on. If the codomain of a random oracle is the field of scalars $\mathbb{F}_p$, we denote this $\mathcal{H}^s$ (hash-to-scalar). If the codomain is $\mathbb{G}$, we denote this $\mathcal{H}^p$ (hash-to-point).

\section{Linkable and $d$-linkable ring signatures}\label{sec:lrs}

In this section, we recall linkable ring signature schemes, related definitions, and relevant examples.

\begin{definition}[Linkable ring signature]\label{def:lrs}
A \textit{linkable ring signature scheme} is a tuple $\left(\texttt{Setup}, \texttt{KeyGen}, \right. \allowbreak \left. \texttt{Sign}, \texttt{Verify}, \texttt{Link}\right)$ satisfying the following.
\begin{itemize}
\item $\texttt{Setup}(1^\lambda) \to \textit{par}$. $\texttt{Setup}$ takes as input a security parameter $1^\lambda$, produces some public parameters $\textit{par}$.

\item $\texttt{KeyGen}(1^\lambda, \textit{par}) \to (sk, pk)$. $\texttt{KeyGen}$ takes as input a security parameter $1^\lambda$ and public parameters $\textit{par}$. $\texttt{KeyGen}$ produces as output a private-public keypair $(sk, pk)$.

\item $\texttt{Sign}\left(1^\lambda, \textit{par}, (m, \textbf{pk}, sk)\right) \to \left\{\bot_{\texttt{Sign}}, \sigma \right\}$. $\texttt{Sign}$ takes as input a security parameter $1^\lambda$, public parameters $\textit{par}$, an arbitrary message $m \in \left\{0,1\right\}^*$, an ad hoc \textit{ring} of public keys $\textbf{pk} = \left\{pk_1, \ldots, pk_n\right\}$, and a secret key $sk$. $\texttt{Sign}$ produces as output either a distinguished failure symbol $\textit{out} = \bot_{\texttt{Sign}}$ or a signature $\sigma$.

\item $\texttt{Verify}\left(1^\lambda, \textit{par}, (m, \textbf{pk}, \sigma)\right) \to \left\{0,1\right\}$. $\texttt{Verify}$ takes as input a security parameter $1^\lambda$, public parameters $\textit{par}$, a message $m$, a ring of public keys $\textbf{pk}$, and a signature $\sigma$. $\texttt{Verify}$ produces as output a bit $b \in \left\{0,1\right\}$; $0$ indicates the signature is not verified, and $1$ indicates the signature is verified.

\item $\texttt{Link}\left(1^\lambda, \textit{par}, (m, \textbf{pk}, \sigma), (m^\prime, \textbf{pk}^\prime, \sigma^\prime)\right)$. $\texttt{Link}$ takes as input a security parameter $1^\lambda$, public parameters $\textit{par}$, and a pair of tuples $(m, \textbf{pk}, \sigma), (m^\prime, \textbf{pk}^\prime, \sigma^\prime)$ for messages $m, m^\prime$, rings $\textbf{pk}$, $\textbf{pk}^\prime$, and signatures $\sigma, \sigma^\prime$. $\texttt{Link}$ produces as output a bit $b \in \left\{0,1\right\}$; $0$ indicates the signatures are not linked or invalid, and $1$ indicates the signatures are valid and linked.\footnote{This is the opposite of the conventions in, say, \cite{tsang2004separable}, which outputs $0$ to indicate two signatures are linked (\textit{i.e.}\ rejected) and outputs $1$ to indicate two signatures are not linked (\textit{i.e.}\ accepted).}
\end{itemize}
\end{definition}

In the sequel, we implicitly assume all algorithms in a LRS take the security parameters as input, and all algorithms (except $\texttt{Setup}$) take public parameters $\textit{par}$ as input. We suppress this notation in the sequel. For example, we write $\textit{par} \leftarrow \texttt{Setup}$ instead of $\textit{par} \leftarrow \texttt{Setup}(1^\lambda)$ and $\sigma \leftarrow \texttt{Sign}(m, \textbf{pk}, \textit{sk})$ instead of $\sigma \leftarrow \texttt{Sign}(1^\lambda, \textit{par}, m, \textbf{pk}, \textit{sk})$.

Note that the vector dimension of keys in Definition \ref{def:lrs} is not specified. If the keys from $\texttt{KeyGen}$ have dimension $d$, we instead say the LRS is a $d$-LRS and use the vector notation introduced in Section \ref{sec:notation} representing keys in boldface (\textit{e.g.}\ \textbf{pk} instead of \textit{pk}) and rings in underlined boldface ($\underline{\textbf{pk}}$ instead of $\textbf{pk}$). We always assume the first coordinate of a key is the \textit{linking key}, and all other coordinates are \textit{auxiliary keys}; we will see that in practice, linkability depends only on the linking coordinate of the signing key vector. For applications in signer-ambiguous confidential transaction models, this is important since only one coordinate in a vector key corresponds to a public key used for double-spend detection purposes.

\subsection{Examples}\label{sec:examples}

In the following examples, all verifiers must list ring public keys in an agreed-upon order for verification to operate properly; in practice, it is simple to use lexicographical ordering.

\begin{example}\label{ex:lsag}
The signature scheme of \cite{liu} is an LRS (and, therefore, a $1$-LRS); we present the variant from \cite{van2013cryptonote} with a linking tag (a signature element which is used for identifying linked signatures) appropriate for linking signatures key-by-key. The signature scheme originally described in \cite{liu} signs a message $m$ with a ring of keys $\textbf{pk} = \left\{\textit{pk}_1, \ldots, \textit{pk}_n\right\}$ and a secret index-key pair $(\ell, \textit{sk})$ corresponding to some $\textit{pk}_\ell$, and the linking tag $\mathfrak{T} := \textit{sk} \cdot \mathcal{H}^p(\textbf{pk})$.

Unfortunately, this tag construction is unsuitable for many applications, as changing ring members will change the linking tag, allowing the same key to sign twice. For use in applications where double-signing detection should be independent of non-signing ring members and following \cite{van2013cryptonote}, we modify the tag from that of \cite{liu} to be $\mathfrak{T} := \textit{sk} \cdot \mathcal{H}^p(\textit{pk}_\ell)$, which is independent of the non-signing ring members. This allows tags to be used for double-spend protection, as discussed below.

$\texttt{Setup}$ always deterministically sets $d := 1$, so there are no auxiliary keys as part of key vectors. $\texttt{Setup}$ selects a generator $G \in \mathbb{G}$ to be a group generator for the group parameters $(p, \mathbb{G}, d, G)$, two cryptographic hash functions $\mathcal{H}^s:\left\{0,1\right\}^* \to \mathbb{F}_p$ and $\mathcal{H}^p:\left\{0,1\right\}^* \to \mathbb{G}$. $\texttt{Setup}$ outputs $\textit{par} = (p, \mathbb{G}, d, G, \mathcal{H}^s, \mathcal{H}^p)$.

$\texttt{KeyGen}$ produces as output a secret key $\textit{sk} \in \mathbb{F}_p^*$ and the corresponding public key $\texttt{pk} = \textit{sk}\cdot G\in \mathbb{G}$.

$\texttt{Sign}$ takes as input a private key $\textit{sk} \in \mathbb{F}_p^*$, a message $m$, a ring $\textbf{pk} = \left\{\textit{pk}_1, \ldots, \textit{pk}_n\right\}$, and produces as output either a distinguished failure symbol $\bot_{\texttt{Sign}}$ or a signature $\sigma$, computed as follows. First, the signer samples field elements $\alpha, s_{\ell+1}, s_{\ell+2}, \ldots, s_{\ell-1} \in \mathbb{F}_p$ at random. Next, the signer computes group elements $H_i = \mathcal{H}^p(\textit{pk}_i)$ and the linking tag $\mathfrak{T} = \textit{sk} \cdot \mathcal{H}^p(\textit{pk}_\ell)$, the first challenge $$c_{\ell+1}  = \mathcal{H}^s(\textbf{pk} \mid \mid m \mid \mid \alpha G \mid \mid \alpha H_\ell)$$ and each subsequent challenge $$c_{i+1} = \mathcal{H}^s(\textbf{pk} \mid \mid m \mid \mid s_i G + c_i \cdot \textit{pk}_i \mid \mid s_i H_i + c_i \mathfrak{T})$$ for $i = \ell+1, \ldots, \ell-1$, naturally identifying index $1$ with index $n+1$. The signer finishes by computing $s_\ell = \alpha - c_\ell \cdot \textit{sk}$ and publishing the signature $\sigma$ where $\sigma = (c_1, s_1, \ldots, s_n, \mathfrak{T})$.

$\texttt{Verify}$ takes as input a message $m$, a ring $\textbf{pk}$, and a purported signature $\sigma^\prime$. $\texttt{Verify}$ parses $(c_1, s_1, \ldots, s_n, \mathfrak{T}) \leftarrow \sigma$. If $c_1 \notin \mathbb{F}_p$ or any $s_i \notin \mathbb{F}_p$ or $\mathfrak{T} \notin \mathbb{G}$, then $\texttt{Verify}$ outputs $0$. Otherwise, the verifier computes $H_i := \mathcal{H}^p(X_i)$ for each ring member, sets $c_1^\prime := c_1$, and computes the challenges $$c_{i+1}^\prime = \mathcal{H}^s(\textbf{pk} \mid \mid m \mid \mid s_i G + c_i^\prime \cdot \textit{pk}_i \mid \mid s_i H_i + c_i^\prime \mathfrak{T})$$ for $i=1, 2, \ldots, n$. The verifier outputs $1$ when $c_{n+1}^\prime = c_1$ and $0$ otherwise.

$\texttt{Link}$ checks the validity of both signatures. If both are valid, $\texttt{Link}$ parses the linking tags $\mathfrak{T}$ and $\mathfrak{T}^\prime$. If either are not in $\mathbb{G}$, the linker outputs $0$ and terminates. Otherwise, the linker outputs $1$ when $\mathfrak{T} = \mathfrak{T}^\prime$ and $0$ otherwise.
\end{example}
